Scalar Feedback based Joint Time-Frequency Precoder interpolation for MIMO-OFDM Systems

### 2020-01-02 paper/main.tex
adding plots with time based of scalar
Minor fixes

Multiple-input multiple-output (MIMO) based wireless communication
systems that use orthogonal frequency division multiplexing (OFDM)
have been adapted in many modern wireless systems owing to their
flexibility and ability to use the wideband channels effectively. One
key requirement to utilize such channels effectively is for the
transmitter to possess channel state information (CSI), since that
permits channel parallelization for low-complexity decoders as well as
the use of capacity achieving transmission
strategies~\cite{love2008overview}. This primarily translates to the
generation of unitary precoding matrices at the transmitter. However,
feeding back CSI from the receiver to the transmitter requires large
feedback overhead, since a precoder needs to be fed back for each OFDM
subcarrier within each coherence interval. Typical past approaches
have used manifold based approaches that exploit temporal and
frequency correlation to reduce feedback overheads. Wireless standards
have largely employed parametric quantization that uses Givens
rotations based quantization~\cite{lou2013comparison,5671092}, though these do
not involve any refinement of the estimates using channel
correlations. In this paper, we propose a joint time-frequency
interpolation approach that builds upon the parametric quantization
techniques used by standards that manages to outperform typical
manifold based quantization techniques.

@@

\begin{figure}
\includegraphics[width=0.4\textwidth]{images/1024timebased}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 3.41\times 10^{-6}$, corresponding to a velocity of 37.8 km/h.}
\label{fig:ber_ped}
\end{figure}

@@

Fig.~\ref{fig:ber_veh} shows a similar comparison for a similar
situation with a higher speed. The faster variation of the channel
taps with time necessitates the use of a smaller number of subcarriers
(viz. 64 subcarriers). The smaller number of subcarriers implies that
the predictive quantization and interpolation needs to be performed
across a larger frequency range, thereby potentially reducing the
accuracy. Even in this case, we observe that the scalar
parameter-based approach outperforms manifold based quantization by
about 1 dB for higher SNRs. Naturally, the rate of channel variation
is quite fast, which limits the efficacy of the proposed method in
tracking the channel parameters. In both the situations, the temporal-only
based interpolation using manifold has the poorest performance since it does
not utilize the joint time and frequency correlations.
\begin{figure}
\includegraphics[width=0.4\textwidth]{images/64timebased}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 1.56\times 10^{-5}$, corresponding to a velocity of 172.8 km/h.}
\label{fig:ber_veh}
\end{figure}

>>>>

Multiple-input multiple-output (MIMO) based wireless communication
systems that use orthogonal frequency division multiplexing (OFDM)
have been adapted in many modern wireless systems owing to their
flexibility and ability to use the wideband channels effectively. To
utilize such channels effectively, the transmitter needs to have the
channel state information (CSI), since this permits channel
parallelization for low-complexity decoders as well as the use of
capacity achieving transmission
strategies~\cite{love2008overview}. This primarily translates to the
generation of unitary precoding matrices at the transmitter. However,
feeding back CSI from the receiver to the transmitter needs large
feedback overheads, since a precoder needs to be fed back for each
OFDM subcarrier within each coherence interval. Past approaches have
typically used manifold based quantization techniques that exploit
temporal and frequency correlation to reduce feedback
overheads. Wireless standards have largely employed parametric
quantization that uses Givens rotations based
quantization~\cite{lou2013comparison,5671092}, though these do not
involve any refinement of the estimates using channel correlations. In
this paper, we propose a joint time-frequency interpolation approach
that builds upon the parametric quantization techniques used by
standards that manages to outperform typical manifold based
quantization techniques.

@@

\begin{figure}
\includegraphics[width=0.4\textwidth]{images/1024final_withtime}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 1024
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 3.41\times 10^{-6}$, corresponding to a velocity of 37.8 km/h.}
\label{fig:ber_ped}
\end{figure}

@@

Fig.~\ref{fig:ber_veh} shows a similar comparison for a similar
situation with a higher speed. The faster variation of the channel
taps with time necessitates the use of a smaller number of subcarriers
(viz. 64 subcarriers). The smaller number of subcarriers implies that
the predictive quantization and interpolation needs to be performed
across a larger frequency range, thereby potentially reducing the
accuracy. Even in this case, we observe that the scalar
parameter-based approach outperforms manifold based quantization by
about 1 dB for higher SNRs. Naturally, the rate of channel variation
is quite fast, which limits the efficacy of the proposed method in
tracking the channel parameters. In both the situations, the temporal-only
based interpolation using manifold has the poorest performance since it does
not utilize the joint time and frequency correlations.
\begin{figure}
\includegraphics[width=0.4\textwidth]{images/64final_withtime}
\caption{BER vs. SNR for QPSK transmission over the $4\times 2$ 64
  subcarrier MIMO-OFDM system, with the ITU Cost207-RA channel profile
  for $f_dT_s = 1.56\times 10^{-5}$, corresponding to a velocity of 172.8 km/h.}
\label{fig:ber_veh}
\end{figure}
